Does personal relevance attenuate citizens' persuasion by government branding? Evidence from a survey experiment on air-pollution policy

### 2019-07-29 paper3_revision_text.Rmd
Add files via upload

*Recent studies have demonstrated the potency of government branding to positively affect citizens' judgments of government organizations and policies. Additionally, studies have pointed to the detrimental implications of the emotive effect of symbols, mainly its ability to compensate for organizations' poor functioning, and accordingly to elicit undue trust. In light of these concerns, this study explores the boundaries of governments' persuasion of citizens through branding and symbolic communications. Building on social psychology and marketing research, I hypothesize that citizens are less susceptible to persuasion by symbolic communication, the more they perceive the policy issue in the communication as personally relevant. I test this expectation through a survey experiment, focused on air-pollution policy in Israel, exploiting the natural variation in the perceived personal relevance between citizens residing in a polluted area in the country and others. The results consistently indicate that even high levels of perceived personal relevance do not significaly attenuate the effect of symbols, which entails that the boundaries of persuasion and manipulation through branding are wider than expected.*

@@

Public Administration scholarship has recently shown a growing interest in public sector branding, and its potency to influence citizens' perceptions and direct their behaviour [@eshuis_2012; @marland_2016; @marland_2017; @zavattaro_website_2018; @sataen_2015]. Recent behavioural studies established that symbolic, brand elements entangled in government communications (e.g. brand names, logos, colors and figures) can have a positive emotional effect on citizens, which leads them to view public organizations and their policies and services more favourably [@marvel_2015a; @karens_2016; @alonbarkat_2017; @teodoro_2018]. This growing scholarly interest corresponds with the overall increase in the use of branding practices by public-sector organizations over the past decade, in part due to the expension of ditigal communication platforms and the use of social media [@eshuis_2012; @marland_2017; @mickoleit_2014; @mergel_2013].

The ability of government organizations to modify citizens' attitudes via such symbolic communications can be perceived favorably, insofar as it mitigates citizens' distrust in organizations and underevaluation of their performance. Accordingly, Symbolic communications, as opposed to mere factual information and substantive arguments, may enable organizations to overcome citizens' negative prior beliefs and biases against the public sector [@marvel_2015a; @marvel_2015; @hvidman_2016; @hvidman_2019]. Nonetheless, this form of persuasion is also highly problematic. Normatively, we expect citizens in a democratic society to form their opinions about their government based on their critical thinking, rather then on uncontious emotive responses. A yet more serious concern regards the potency of symbolic communication to compensate for organizations' poor performance, poorly planned policy plans and logically unpersuasive explanations, as indicated by previous studies [@alonbarkat_2017] AUTHOR. Congruently, government branding can yield a misalignment between citizens' positive view and the "real world", which would undermine the democratic accountability and responsiveness of public organizations. Moreover, government agencies and/or politicians controlling them could use branding symbolic communication strategically as an instrument of propaganda, seeking to manipulate public opinion and mitigate justified public criticism of agencies' failures.

The abovementioned concerns of the negative implications of branding for democracy are the main motivation of this study. These concerns necessitate a better theoretical understanding of the boundaries of governments' persuasion (and possibly manipulation) of citizens through symbolic communication, which would enable to asses more effectively the scope of the risk that these practices pose for democracy. In line with this motivation, this study focuses of an importnant potential limitation that regards the perceived personal relevance of the policy issue. I explore whether the effect of symbolic elements in communications is restricted to policy issues that are not perceived by citizens as having a significant consequence for their personal lives, or rather that citizens are similarly susceptible to persuasion via branding even with regard to such personally relevant issues. This factor is particularly important, since we tend to expect these affected groups to pay more attention to government actions that concern their interests, and to hold them to account for their consequences to a greater extent. That expectation can be supported, theoretically, by the social psychology theory Elaboration Likelihood Model (ELM) [@petty_1986; @petty_2011]. The model postulates that the more people perceive the communication as personally relevant, the more inclined they are to scrutinize its message, and accordingly the less likely they are to rely on peripheral, symbolic elements.

Employing a survey experiment, I test this latter theoretical expectation in the context of the Israeli Environment Protection Ministry, and its policy plan for reducing the air pollution in a specific geographical area -- the Haifa Bay. The research sample (*n* = `r SVIVA2_01 %>% nrow()`) consists of citizens residing in the polluted area, matched with residents of other cities in the center of the country that are not exposed to air-pollution. I present all participants a policy plan regarding the air-pollution in the Haifa Bay area, while experimentally varying both its appearance (familiar symbolic elements from the Ministry's communications versus two control groups) and content (logically persuasive versus unpersuasive policy plans). I examine the effects of these manipulations on participants' trust in the policy plan accross the two areas, and test whether residents of the polluted area are less affected by familiar symbolic elements (and more affected by the substantive differences). I further supplement this observational comparison between the areas with an experimental treatment for the perceived personal relevance of the policy, and with additional robustness tests.

Contrary to my initial expectations, the results indicate that even high levels of perceived personal relevance do not significantly attenuate the effect of symbols on citizens' trust. These findings tend to suggest that the boundaries of the effect of symbolic communication are wider than expected. I discuss the theoretical and normative implications of these findings.

@@

A brand can be defined as *"a symbolic construct that consists of a name, term, sign, symbol, or design, or combination of these, created deliberately to identify and phenomenon and differentiate it from similar phenomena by adding particular meaning to it"* [@eshuis_2012, p. 19]. Branding, therefore, can be considerred a form of symbolic communication. In practice, symbolic elements that are strategically selected and designed to evoke positive emotions and associations, can be found in almost every government communication with citizens.

@@

According to the Elaboration Likelihood Model (ELM) [@petty_1986; @petty_2011], persuasion can occur via different process that can be placed along an "elaboration continuum". At the high end of that continuum, also called "central route", persuasion occurs through processes that require high degrees of thought and cognitive effort. This type of processing involves thoughtful scrutiny of arguments in the communication that are central to the merits of the issue. People located at the high end tend are more likely to form their judgements on the matter based on the quality of the arguments. They are more likely to differentiate between strong and weak arguments, and would be persuaded mainly by strong, logically persuasive arguments. As for those who are located at the low end of that continuum, also called "peripheral route", persuasion may occur mainly through processes that require little ammount of thinking, such as evaluative conditioning. Another example for peripheral route processing is people's reliance on the message source as a heuristic cue [@james_2017a]. When people are located between the two ends of the continuum, a mixture of these persuasion processes will operate.

In any given situation, the position of people along that continuum (and accordingly the dominant type of processing) would be determined by their motivation and ability to scrutinize the message. The more motivated people are to elaborate on the message and the more capable they are of doing so, the more likely they are to focus on the substative arguments, and likewise the less susceptible they are to persuasion via peripheral route processes. ELM literature has pointed to a variety of factors that may shape people's motivation and ability to think about a message, and accordingly affect their inclination to be persuaded either by the arguments or by peripheral cues. Perhaps the most studied factor in this regard is the perceived personal relevance of the communication (or "issue involvement"), namely the extent to which they perceive the issue at hand as having significant consequences for their own lives. That variable has been considered the most important determinant of motivation to scrutinize the message [@petty_2002; @petty_2011].[^footnote_elm_factors] It has been theorized that the more people perceive an issue as personally relevant, the more motivated they are to invest in thinking about it, since the consequences of being incorrect are greater [@petty_1979; @petty_1981; @petty_1983; @johnson_1989; @petty_1990; @burnkrant_1989]. Accordingly, experimental studies have showed that individuals are more affected by peripheral cues and less affected by argument quality, the less they perceive the issue as personally relevant.

For instance, in a study by @petty_1983, participants were asked to express their attitudes about a new product (disposable razor), after being exposed to a magazine ad. The ad involved either a celebrity endorser or a non-famous endorser (a symbolic peripheral cue) and contained substantive information about the product that represent either strong or weak arguments about its quality. In addition, scholars manipulated the perceived personal relevance of the product by telling subjects either that the product would soon be available in their local area (high relevance), or that it would be available in other areas (low relevance). The results of the study indicated that the celebrity endorser had a positive effect on participants' attitudes about the product under the low personal relevance condition (estimated as roughly 0.7 SDs), but had no effect under the high personal relevance condition. Likewise, the effect of argument quality was greater for the high personal relevance group (1.95 SDs versus 0.54 SDs in the low personal relevance group).

@@

I was assisted by an Israeli internet research panel company *iPanel* for the selection and recruitment of residents of these selected cities. The survey experiment was conducted online through Qualtrics Survey Software. The online survey link was sent by iPanel between 28.1-1.2.2018. A total of `r N_0` respondents completed the survey (a response rate of `r round((N_0/5077)*100,1)`%). Thereafter, I filtered out `r N_0-N_1` observations (`r round(((N_0-N_1)/N_0)*100,1)`%) due to following criteria: multiple entries from the same IP address (n=`r filter_IP`), failure at the instructional manipulation test (n=`r filter_IMC`), submitting the survey in less than 3 or more than 30 minutes (n=`r filter_time`), and reported age under 18 (n=`r filter_age`). These filters are neither associated with the areas nor with the experimental conditions. Following this screening, I ended up with a sizable analytical sample of `r N_1` subjects, of whom `r SVIVA2_01_haifa%>%nrow()` live in the Haifa Bay cities, and `r SVIVA2_01_center%>%nrow()` live in the center cities.[^footnote_unfiltered_sample]

@@

The survey was presented to the participants as an academic study about citizens' attitudes regarding environmental policies in general. After a set of pre-manipulation questions, all subjects, regardless of their area of residence, were randomly assigned to treatment and control conditions of perceived personal relevance (hereafter: relevance manipulation). Subjects in the relevance treatment group were asked three questions about their interest in environmental policy issues, the extent to which they follow these issues in the media, and their preference for more/less government investment in this field. Then, they were asked to describe the most acute environmental issue in their area of residence. These treatment questions were designed to prime those residing in the Haifa Bay area to think about the personal relevance of air-pollution in their area.[^footnote_relevance_manipulation] Subjects in the relevance control group were asked equivalent questions about their field of occupation.

[^footnote_relevance_manipulation]: The first three treatment questions were also initially designed to prime subjects, more generally, to think about the impact of environmental issues to their own lives. In this paper, I focus on the the moderating effect of the relevance manipulation on Haifa Bay citizens' responses to the Haifa Bay air-pollution policy.

Participants were then presented with the air-pollution and recycling policy plans (displayed in a random order). They were informed that these plans were obtained from EPM's annual work plan. The participants were randomly assigned to one of three conditions of symbolic elements in the communication (hereafter: symbols manipulation): treatment ("real symbols"), control-"no symbols" and control-"fake symbols". In the real symbols condition, the two policies are displayed with three symbolic brand elements, which have been used in EPM's public communications, and became widely recognized by the public. The first symbolic element is the unique brand logo of the EPM, which consists a pair of green and orange leaves that resemble two hands, designed to symbolize peace and harmony with the environment. The second element is the green color, which is strongly associated with taking care of the environment. The third element is an image of a celebrity comedian from an advertising campaign of the EPM. The Haifa Bay air-pollution policy (in the real symbols condition) was presented with an image from EPM campaign from 2010 under the slogan "starting to think green", starring the Israeli celebrity comedian Tal Friedman. The recycling policy (in the real symbols condition) was presented with an image from a campaign from 2017 for reducing the use of disposable bags ("taking every bag seriously"), starring the comedian Ido Rosenblum. In the control no-symbols condition, the two policies are displayed in a minimal black and white design, and without the logo and the celebrities’ images. Finally, in the control fake-symbols condition, the EPM logo is replaced by a fake logo, the green color is replaced with blue, and the images of celebrities from the campaigns are replaced with edited images of unfamiliar people. The purpose of the fake-logo condition was to distinguish the effect of the symbolic meaning of the EPM's graphic brand elements (i.e. the symbolic effect) from a possible effect of their aesthetic design. Accordingly, I specifically designed the "fake" symbols so that they would resemble the real symbols and have equivalent aesthetic qualities, without activating particular emotions and associations related to EPM, or environmental policy. The symbols conditions are displayed in **Appendix B**. Further details about the real and fake symbols are found the supplementary appendix (section 2).

In addition, I also included a manipulation of the argument quality of the information in the communication about the policy plan (hereafter: information manipulation). Per each policy, subjects were randomly allocated to one of two conditions of information: strong policy (i.e. logically persuasive policy plan) or weak policy (i.e. logically unpersuasive). Each subject was exposed to two policy plans, one strong and one weak, displayed in a random order. For instance, a subject who first received a strong Haifa Bay air-pollution policy plan, later received the weak recycling policy plan, and vice versa. The information in the strong policy plan condition was taken from the EPM’s real work plan and from other official sources, whereas the information in the weak policy plan was fictional, and deliberately designed to create a poor policy plan, which is incompatible with achieving the policy goals.[^footnote_reliability_information_manipulation] The full texts of the weak and strong conditions for the two policy plans is displayed in **Appendix A**.

@@

Living in Haifa Bay area was coded according to respondents’ reporting of residence, work or studying in one of the Haifa Bay cities. To confirm the empirical assumption that the Haifa Bay air-pollution policy (but not the recycling policy) is perceived as more personally relevant by Haifa Bay residents compared with center residents, respondents were asked, after the outcome variable items: "To what extent did you feel that the policy on ["Reducing the air-pollution in the Haifa Bay" / “reducing waste and increasing recycling”] is personally relevant to you?" (1="very little"; 7="very much"). `r figure.x=figure.x+1`**Figure `r figure.x`** shows the distributions of the latter items across the two areas. As expected, the Haifa Bay air-pollution policy is perceived as more personally relevant by Haifa Bay residents compared with center residents  (Cohen's D = `r relevance.areas.d[["estimate"]] %>% round(2)` [`r relevance.areas.d[["conf.int"]][["lower"]] %>% round(2)`,`r relevance.areas.d[["conf.int"]][["upper"]] %>% round(2)`]). By comparison, the recycling policy is perceived by both groups as relatively personally relevant by both groups, with no statistically significant differences between them. Additional support for these differences are provided by respondents' answers to the open question about the most salient environmental issue in their area of residence, which was presented to the relevance treatment group (i.e. before presenting the policy plans). Indeed, `r t1<-SVIVA2_01_haifa %>% filter(RELEVANCE_exp==1,air.content==1) %>% nrow()``r t1` of the `r t2<-SVIVA2_01_haifa %>% filter(RELEVANCE_exp==1) %>% nrow()``r t2` Haifa Bay subjects who were asked this question regarded in their responses to the air-pollution generated from the industrial area (`r air.content.percent.haifa <- t1/t2*100``r air.content.percent.haifa %>% round(0)`%).[^footnote_air_relevance_coding] Moreover, many Haifa Bay respondents explicitly linked the air-pollution to higher morbidity rates in their area. Subjects from center cities regarded to a variety of issues, including recycling, traffic jams and cleanliness, and only `r ((SVIVA2_01_center %>% filter(RELEVANCE_exp==1,air.content==1) %>% nrow())/nrow(SVIVA2_01_center)*100) %>% round(0)`% of them mentioned air-pollution as a prominent environmental concern in their area.

@@

As for the relevance manipulation, data suggests that assignment to relevance priming questions did not increase Haifa Bay respondents' perceptions of personal relevance of the air-pollution policy. In fact, those assigned to the treatment group reported sightly lower level of personal relevance on average, yet these differences are not statistically significant (*t* = `r relevance.exp.t[["statistic"]] %>% round(3)`, *p*=`r round(relevance.exp.t[["p.value"]],3)`). This limitation should be kept in mind when interpreting the results regarding this manipulation. The results of all manipulation and validation checks are available from the supplementary appendix, along with summary statistics and correlation matrix for all variables (sections 3, 5).

@@

[^footnote_raw_means]: Means and SDs of outcome variables across the two areas and the experimental conditions are reported in section XX of the supplementary appendix (section 6).

@@

I further examined the interactions between the manipulations and the two areas via a series of multiple linear regression models, presented in `r table.x=table.x+1`**Table `r table.x`**. I regress trust in Haifa Bay air-pollution policy on the manipulations (Model `r table.x`.1), then I add the areas (Model `r table.x`.2) and the interaction between them (Model `r table.x`.3). In Model `r table.x`.4, I test the robustness of the interaction model by controlling for individual-level covariates.[^footnote_control_variables] In all regression models, I use robust standard errors to adjust for heteroskedacity.[^footnote_heteroskedacity] The results of these models are largely consistent with the raw findings presented above. In Models `r table.x`.3, `r table.x`.4 the interactions between living in the center and the real symbols are negative but not sufficiently significant, and the interactions with the weak information are insignificant. A comparison between models `r table.x`.2 and `r table.x`.3 via anova reveals that adding these interactions do not significantly improve the model fit (*F(`r anova(mod_air_1.2,mod_air_1.3) %>% tidy() %>% select(df) %>% .[2,] %>% max()`*) = `r anova(mod_air_1.2,mod_air_1.3) %>% tidy() %>% select(statistic) %>% .[2,] %>% max() %>% round(2)`, *p*= `r anova(mod_air_1.2,mod_air_1.3) %>% tidy() %>% select(p.value) %>% .[2,] %>% max() %>% round(2)`).

@@

Overall, the findings with regard to the Haifa Bay air-pollution policy indicate, tentatively, that higher levels of perceived personal relevance of the communication neither attenuate the positive effect of brand elements entangled in communications, nor enhances citizens' attention to their substantive content. Still, as discussed in the previous section, this analysis is based on the observational comparison between the areas and not by a random assignment, and hence a concern of endogeneity cannot be entirely ruled. Accordingly, the null findings regarding the interactions between the areas and the manipulations can be resulted by other, unobserved differences between the two populations. Yet, as explained, the experimental design provides two additional tests which enable to address this concern: the comparison to the recycling policy, for which there are no relevance differences between the two groups, and the relevance manipulation. I now discuss the results of these two tests.

>>>>

*Recent studies have demonstrated the potency of government branding to positively affect citizens' judgments of government organizations and policies. Additionally, studies have pointed to the detrimental implications of the emotive effect of symbols, mainly its ability to compensate for organizations' poor functioning, and accordingly to elicit undue trust. In light of these concerns, this study explores the boundaries of governments' persuasion of citizens through branding and symbolic communications. Building on social psychology and marketing research, I hypothesize that citizens are less susceptible to persuasion by symbolic communication, the more they perceive the policy issue in the communication as personally relevant. I test this expectation through a survey experiment, focused on air-pollution policy in Israel, exploiting the natural variation in the perceived personal relevance between citizens residing in a polluted area in the country and others. The results consistently indicate that even high levels of perceived personal relevance do not significantly attenuate the effect of symbols, which entails that the boundaries of persuasion and manipulation through branding are wider than expected.*

@@

Public Administration scholarship has shown a growing interest in public sector branding, and its potency to influence citizens' perceptions and direct their behaviour [@eshuis_2012; @marland_2016; @marland_2017; @zavattaro_website_2018; @sataen_2015]. Congruently, recent behavioural studies established that symbolic, brand elements entangled in government communications (e.g. brand names, logos, colors and figures) can have a positive emotional effect on citizens, which leads them to view public organizations and their policies and services more favourably [@marvel_2015a; @karens_2016; @alonbarkat_2017; @teodoro_2018]. This growing scholarly interest corresponds with the overall increase in the use of branding practices by public-sector organizations over the past decade or so, in part due to the vast expansion of ditigal communication and social media [@eshuis_2012; @marland_2017; @mickoleit_2014; @mergel_2013].

The ability of government organizations to modify citizens' attitudes via such symbolic communications can be perceived favorably, insofar as it mitigates citizens' distrust in organizations and underevaluation of their performance. Accordingly, symbolic communications, as opposed to the delivery of mere factual information and substantive arguments, may enable organizations to overcome citizens' negative prior beliefs and biases against the public sector [@marvel_2015a; @marvel_2015; @hvidman_2016; @hvidman_2019]. Nonetheless, this form of persuasion is also highly problematic. Normatively, we expect citizens in a democratic society to form their opinions about their government based on their critical thinking, rather then on uncontious emotive responses. A yet more serious concern regards the potency of symbolic communication to compensate for organizations' poor performance, poorly planned policy plans and logically unpersuasive explanations, as indicated by previous studies [@alonbarkat_2017, AUTHOR]. Congruently, government branding can yield a misalignment between citizens' positive view and the "real world", which would undermine governments' democratic accountability and responsiveness. Moreover, government agencies and/or politicians controlling them could use branding and symbolic communication strategically as an instrument of propaganda, seeking to manipulate public opinion and mitigate justified public criticism of agencies' failures.

The abovementioned concerns of the negative implications of branding for democracy are the main motivation of this study. These concerns necessitate a better theoretical understanding of the boundaries of governments' persuasion (and possibly manipulation) of citizens through symbolic communication, which would enable us to asses more effectively the scope of the risk that these practices pose for democracy. In line with this motivation, this study focuses of an importnant potential limitation that regards the perceived personal relevance of the policy issue. I explore whether the effect of symbolic elements in communications is restricted to policy issues that are not perceived by citizens as having a significant consequence for their personal lives, or rather that citizens are similarly susceptible to persuasion via branding even with regard to such personally relevant issues. This moderating factor is particularly important, since we tend to expect these affected groups to pay more attention to government actions that concern their interests, and to hold them to account for their consequences to a greater extent. That expectation can be supported, theoretically, by the social psychology theory Elaboration Likelihood Model (ELM) [@petty_1986; @petty_2011]. The model postulates that the more people perceive the communication as personally relevant, the more inclined they are to scrutinize its message, and accordingly the less likely they are to rely on peripheral, symbolic elements.

Employing a survey experiment, I test this latter theoretical expectation in the context of the Israeli Environment Protection Ministry, and its policy plan for reducing the air pollution in a specific geographical area -- the Haifa Bay. The research sample (*n* = `r SVIVA2_01 %>% nrow()`) consists of citizens residing in the polluted area, matched with residents of other cities in the center of the country that are not exposed to high levels of air-pollution. I present all participants a policy plan regarding the air-pollution in the Haifa Bay area, while experimentally varying both its appearance (familiar symbolic elements from the Ministry's communications versus two control groups) and substantive information (logically persuasive versus unpersuasive policy plans). I examine the effects of these manipulations on participants' trust in the policy plan accross the two areas, and test whether residents of the polluted area are less affected by familiar symbolic elements (and more affected by the differences in substantive arguments). I further supplement this observational comparison between the areas with an experimental treatment for the perceived personal relevance of the policy, and with additional robustness tests.

Contrary to my initial expectations, the results consistently indicate that even high levels of perceived personal relevance do not significantly attenuate the effect of symbols on citizens' trust. These findings tend to suggest that the boundaries of the effect of symbolic communication are wider than expected. I discuss the theoretical and normative implications of these findings.

@@

A brand can be defined as *"a symbolic construct that consists of a name, term, sign, symbol, or design, or combination of these, created deliberately to identify and phenomenon and differentiate it from similar phenomena by adding particular meaning to it"* [@eshuis_2012, p. 19]. Branding, therefore, can be considerred a form of symbolic communication. In practice, symbolic elements that are strategically designed to evoke positive emotions and associations, can be found in almost every government communication with citizens.

@@

According to the Elaboration Likelihood Model (ELM) [@petty_1986; @petty_2011], persuasion can occur via different process that can be placed along an "elaboration continuum". At the high end of that continuum, also called "central route", persuasion occurs through processes that require high degrees of thought and cognitive effort. This type of processing involves thoughtful scrutiny of arguments in the communication that are central to the merits of the issue. People located at the high end tend are more likely to form their judgements on the matter based on the quality of the arguments. They are more likely to differentiate between strong and weak arguments, and would be persuaded mainly by strong, logically persuasive arguments. As for those who are located at the low end of that continuum, also called "peripheral route", persuasion may occur mainly through processes that require little ammount of thinking, such as evaluative conditioning. Another prominent example for peripheral route processing is people's reliance on the message source as a heuristic cue [@james_2017a]. When people are located between the two ends of the continuum, a mixture of these persuasion processes will operate.

In any given situation, the position of people along that continuum (and accordingly the dominant type of processing) would be determined by their motivation and ability to scrutinize the message. The more motivated people are to elaborate on the message and the more capable they are of doing so, the more likely they are to focus on the substative arguments, and likewise the less susceptible they are to persuasion via peripheral route processes. ELM literature has pointed to a variety of factors that may shape people's motivation and ability to think about a message, and accordingly affect their inclination to be persuaded either by the arguments or by peripheral cues. Perhaps the most studied factor in this regard is the perceived personal relevance of the communication (or "issue involvement"), namely the extent to which they perceive the issue at hand as having significant consequences for their own lives. That variable has been considered the most important determinant of motivation to scrutinize the message [@petty_2002; @petty_2011].[^footnote_elm_factors] It has been theorized that the more people perceive an issue as personally relevant, the more motivated they are to invest in thinking about it, since the consequences of being incorrect are greater [@petty_1979; @petty_1981; @petty_1983; @johnson_1989; @petty_1990; @burnkrant_1989]. Accordingly, experimental studies in marketing and social psychology have showed that individuals are more affected by peripheral cues and less affected by argument quality, the less they perceive the issue as personally relevant.

For instance, in a study by @petty_1983, participants were asked to express their attitudes about a new product (disposable razor), after being exposed to a magazine ad. The ad involved either a celebrity endorser or a non-famous endorser (a symbolic peripheral cue) and contained substantive information about the product that represent either strong or weak arguments about its quality. In addition, scholars manipulated the perceived personal relevance of the product by telling subjects either that the product would soon be available in their local area (high relevance), or that it would be available in other areas (low relevance). The results of the study indicated that the celebrity endorser had a positive effect on participants' attitudes about the product under the low personal relevance condition (estimated as roughly 0.7 SDs), but had no effect under the high personal relevance condition. Likewise, the effect of argument quality was more than three times greater for the high personal relevance group.

@@

I was assisted by an Israeli internet research panel company *iPanel* for the selection and recruitment of residents of these selected cities. The survey experiment was conducted online through Qualtrics Survey Software. The online survey link was sent by iPanel between 28.1-1.2.2018. A total of `r N_0` respondents completed the survey (a response rate of `r round((N_0/5077)*100,1)`%). I further filtered out `r N_0-N_1` observations (`r round(((N_0-N_1)/N_0)*100,1)`%) due to following criteria: multiple entries from the same IP address (n=`r filter_IP`), failure at the instructional manipulation test (n=`r filter_IMC`), submitting the survey in less than 3 or more than 30 minutes (n=`r filter_time`), and reported age under 18 (n=`r filter_age`). These filters are neither associated with the areas nor with the experimental conditions. Following this screening, I ended up with a sizable analytical sample of `r N_1` subjects, of whom `r SVIVA2_01_haifa%>%nrow()` live in the Haifa Bay cities, and `r SVIVA2_01_center%>%nrow()` live in the center cities.[^footnote_unfiltered_sample]

@@

The survey was presented to the participants as an academic study about citizens' attitudes regarding environmental policies in general. After a set of pre-manipulation questions, all subjects, regardless of their area of residence, were randomly assigned to treatment and control conditions of perceived personal relevance (hereafter: relevance manipulation). Subjects in the relevance treatment group were asked three questions about their perceptions of environmental policy in general. They were then asked to state their area  and locaity of residence, and to describe, in their own words, the most disturbing environmental issue in their area. These questions were designed to prime those residing in the Haifa Bay area to think about the personal relevance of air-pollution in their area.[^footnote_relevance_manipulation] Subjects in the relevance control group were asked equivalent questions about their field of occupation.

[^footnote_relevance_manipulation]: The first three questions are: "To what extent are you interested in environmental issues?"; "To what extent do you follow environmental issues in the media and/or in social networks?"; "Compared to the current situation, do you think the government should invest more or less in dealing with environmental issues?" These questions were initially designed to prime participants to think not only about the relevance of the Haifa Bay air-pollution, but of environmental issues general. Yet, I focus here on the moderation of the relevance manipulation on Haifa Bay citizens' responses to the air-pollution policy.

Participants were then presented with the two policy plans (Haifa Bay air-pollution and recycling), displayed in a random order. They were informed that these plans were obtained from EPM's annual work plan for the following year. The participants were randomly assigned to one of three conditions of symbolic elements in the communication (hereafter: symbols manipulation): treatment ("real symbols"), control-"no symbols" and control-"fake symbols". In the real symbols condition, the two policies are displayed with three symbolic brand elements, which have been used in EPM's public communications, and became widely recognized by the public. The first symbolic element is the unique brand logo of the EPM, which consists a pair of green and orange leaves that resemble two hands, designed to symbolize peace and harmony with the environment. The second element is the green color, which is strongly associated with taking care of the environment. The third element is the images of celebrity comedians from advertising campaigns of the EPM. The Haifa Bay air-pollution policy (in the real symbols condition) was presented with an image from EPM campaign from 2010 under the slogan "starting to think green", starring the Israeli celebrity comedian Tal Friedman. The recycling policy (in the real symbols condition) was presented with an image from a campaign from 2017 for reducing the use of disposable bags ("taking every bag seriously"), starring the comedian Ido Rosenblum. In the control no-symbols condition, the two policies are displayed in a minimal black and white design, and without the logo and the celebrities’ images. Finally, in the control fake-symbols condition, the EPM logo is replaced by a fake logo, the green color is replaced with blue, and the images of celebrities from the campaigns are replaced with edited images of unfamiliar people. The purpose of the fake-logo condition was to distinguish the effect of the symbolic meaning of the EPM's graphic brand elements (i.e. the symbolic effect) from a possible effect of their aesthetic design. Accordingly, I specifically designed the "fake" symbols so that they would resemble the real symbols and have equivalent aesthetic qualities, without activating particular emotions and associations related to EPM, or environmental policy. Hence, these elements can be also regarded as peripheral cues (according to ELM conceptualization), yet they do not represent familiar symbols of EPM. The symbols conditions are displayed in **Appendix A**. Further details about the real and fake symbols are found the supplementary appendix (section 2).

In addition, I also included a manipulation of the argument quality of the information in the communication about the policy plan (hereafter: information manipulation). Per each policy, subjects were randomly allocated to one of two conditions of information: strong policy (i.e. logically persuasive policy plan) or weak policy (i.e. logically unpersuasive). Each subject was exposed to two policy plans, one strong and one weak, displayed in a random order. For instance, a subject who first received a strong Haifa Bay air-pollution policy plan, later received the weak recycling policy plan, and vice versa. The information in the strong policy plan condition was taken from the EPM’s real work plan and from other official sources, whereas the information in the weak policy plan was fictional, and deliberately designed to create a poor policy plan, which is incompatible with achieving the policy goals.[^footnote_reliability_information_manipulation] The full texts of the weak and strong conditions for the two policy plans is displayed in **Appendix B**.

@@

Living in Haifa Bay area was coded according to respondents’ reporting of residence, work or studying in one of the Haifa Bay cities. To confirm the empirical assumption that the Haifa Bay air-pollution policy (but not the recycling policy) is perceived as more personally relevant by Haifa Bay residents compared with center residents, respondents were asked, after the outcome variable items: "To what extent did you feel that the policy on ["Reducing the air-pollution in the Haifa Bay" / “reducing waste and increasing recycling”] is personally relevant to you?" (1="very little"; 7="very much"). `r figure.x=figure.x+1`**Figure `r figure.x`** shows the distributions of the latter items across the two areas. As expected, the Haifa Bay air-pollution policy is perceived as more personally relevant by the local residents compared with those residing in the center (Cohen's D = `r relevance.areas.d[["estimate"]] %>% round(2)` [`r relevance.areas.d[["conf.int"]][["lower"]] %>% round(2)`,`r relevance.areas.d[["conf.int"]][["upper"]] %>% round(2)`]). By comparison, the recycling policy is perceived by both groups as relatively personally relevant by both groups, with no statistically significant differences between them. Additional support for these differences are provided by respondents' answers to the open question presented to the relevance treatment group about the most salient environmental issue in their area of residence. Indeed, `r t1<-SVIVA2_01_haifa %>% filter(RELEVANCE_exp==1,air.content==1) %>% nrow()``r t1` of the `r t2<-SVIVA2_01_haifa %>% filter(RELEVANCE_exp==1) %>% nrow()``r t2` Haifa Bay subjects who were asked this question regarded in their responses to the air-pollution generated from the industrial area (`r air.content.percent.haifa <- t1/t2*100``r air.content.percent.haifa %>% round(0)`%),[^footnote_air_relevance_coding] and many of them explicitly linked it to higher morbidity rates in their area. Subjects from center cities regarded to a variety of issues, including recycling, traffic jams and cleanliness, and only `r ((SVIVA2_01_center %>% filter(RELEVANCE_exp==1,air.content==1) %>% nrow())/nrow(SVIVA2_01_center)*100) %>% round(0)`% of them mentioned air-pollution as a prominent environmental concern in their area.

@@

As for the relevance manipulation, data suggests that assignment to relevance priming questions did not increase Haifa Bay respondents' perceptions of personal relevance of the air-pollution policy. The Haifa Bay participants assigned to the treatment group reported sightly lower level of personal relevance on average, yet these differences are not statistically significant (Cohen's D = `r relevance.exp.d[["estimate"]] %>% round(2)` [`r relevance.exp.d[["conf.int"]][["lower"]] %>% round(2)`,`r relevance.exp.d[["conf.int"]][["upper"]] %>% round(2)`], *t* = `r relevance.exp.t[["statistic"]] %>% round(3)`, *p*=`r round(relevance.exp.t[["p.value"]],3)`). This limitation should be kept in mind when interpreting the results regarding this manipulation. The results of all manipulation and validation checks are available from the supplementary appendix, along with summary statistics and correlation matrix for all variables (sections 3, 5).

@@

[^footnote_raw_means]: Means and SDs of outcome variables across the two areas and the experimental conditions are reported in the supplementary appendix (section 6).

@@

I further examined the interactions between the manipulations and the two areas via a series of multiple linear regression models, presented in `r table.x=table.x+1`**Table `r table.x`**. I regress trust in Haifa Bay air-pollution policy on the manipulations (Model 1.1), then I add the areas (Model 1.2) and the interaction between them (Model 1.3). In Model 1.4, I test the robustness of the interaction model by controlling for individual-level covariates.[^footnote_control_variables] In all regression models, I use robust standard errors to adjust for heteroskedacity.[^footnote_heteroskedacity] The results of these models are largely consistent with the raw findings presented above. In Models 1.3, 1.4 the interactions between living in the center and the real symbols are negative but and significant at the 90% level, and the interactions with the weak information are insignificant. A comparison between models 1.2 and 1.3 via anova reveals that adding these interactions do not significantly improve the model fit (*F(`r anova(mod_air_1.2,mod_air_1.3) %>% tidy() %>% select(df) %>% .[2,] %>% max()`*) = `r anova(mod_air_1.2,mod_air_1.3) %>% tidy() %>% select(statistic) %>% .[2,] %>% max() %>% round(2)`, *p*= `r anova(mod_air_1.2,mod_air_1.3) %>% tidy() %>% select(p.value) %>% .[2,] %>% max() %>% round(2)`).

@@

In summary, the findings with regard to the Haifa Bay air-pollution policy indicate, tentatively, that higher levels of perceived personal relevance of the communication neither attenuate the positive effect of brand elements entangled in communications, nor enhances citizens' attention to their substantive content. Still, as discussed in the methodology section, this analysis is based on an observational comparison, and hence a concern of endogeneity cannot be entirely ruled. Accordingly, the null findings regarding the interactions between the areas and the manipulations can be resulted by other, unobserved differences between the two populations. Yet, as explained, the experimental design provides two additional tests which enable to address this concern: the comparison to the recycling policy, for which there are no relevance differences between the two areas, and the relevance manipulation. I now discuss the results of these two tests.
